---
title: "Open Scholarly Data @ SUB Göttingen"
format:
  html:
    page-layout: full
toc-title: Datasets
params:
  bq_project: "subugoe-collaborative"
---

The Scholarly Communication Analytics team at the State and University Library in Göttingen maintains a publicly accessible Open Scholarly Data Warehouse, which is based on Google BigQuery. 

The warehouse features monthly Crossref snapshots, as well as data from various other sources, including OpenAlex, Semantic Scholar and Unpaywall, and provides access to bibliometric data from the German Competence Network for Bibliometrics.

Google BigQuery is provided as part of the OCRE 2024 Framework, with support from the GWGD.

More info: <https://subugoe.github.io/scholcomm_analytics/>

Contact: Najko Jahn

```{r}
#| echo: false
#| message: false
library(tidyverse)
library(bigrquery)
library(reactable)
library(htmltools)

# Write credentials to temp file if BQ_ORI contains JSON content
bq_ori <- Sys.getenv("BQ_ORI")
if (bq_ori != "") {
  if (file.exists(bq_ori)) {
    # BQ_ORI is a file path
    bigrquery::bq_auth(path = bq_ori)
  } else {
    # BQ_ORI contains JSON content (from GitHub Actions)
    temp_key <- tempfile(fileext = ".json")
    cat(bq_ori, file = temp_key)
    bigrquery::bq_auth(path = temp_key)
  }
}
```

```{r}
#| echo: false
my_datasets <- bq_project_datasets(params$bq_project)

# Function to extract nested fields from RECORD type
get_nested_fields <- function(field) {
  if (field$type == "RECORD" && !is.null(field$fields)) {
    tibble::tibble(
      Field = map_chr(field$fields, "name"),
      Type = map_chr(field$fields, "type"),
      Description = map_chr(field$fields, \(x) x$description %||% NA_character_)
    )
  } else {
    NULL
  }
}
```

```{r}
#| echo: false
#| results: asis

# Render a section for each dataset
for (dataset in my_datasets) {
  # Print dataset name as heading
  cat(paste0("## ", dataset$dataset, "\n\n"))

  # Get dataset metadata
  dataset_meta <- bq_dataset_meta(dataset)

  # Create HTML for dataset metadata with styled box
  dataset_url <- paste0(
    'https://console.cloud.google.com/bigquery?project=', dataset$project,
    '&p=', dataset$project,
    '&d=', dataset$dataset,
    '&page=dataset'
  )

  dataset_info_html <- paste0(
    '<div class="description-box">',
    if (!is.null(dataset_meta$description) && dataset_meta$description != "") {
      paste0(
        '<div class="description-header">Description</div>',
        '<div class="description-content" style="margin-bottom: 0.75rem;">',
        dataset_meta$description,
        '</div>'
      )
    } else "",
    '<div class="description-meta">',
    '<strong>Created:</strong> ',
    strftime(as.POSIXct(as.numeric(dataset_meta$creationTime) / 1000, origin = "1970-01-01"),
             format = "%b %d, %Y %H:%M"),
    ' | <strong>Location:</strong> ', dataset_meta$location,
    ' | <a href="', dataset_url, '" target="_blank">View in BigQuery Console</a>',
    '</div>',
    '</div>\n\n'
  )

  cat(dataset_info_html)

  # Set the current dataset for the child document
  current_dataset <- dataset

  # Render the child document
  res <- knitr::knit_child("../_dataset_table.qmd", quiet = TRUE, envir = environment())
  cat(res, sep = "\n")
  cat("\n\n")
}
```
