---
title : Collections Overview
format:
  html:
    toc: false
---

```{r}
#| echo: false
#| message: false
library(tidyverse)
library(bslib)

stats <- read_csv(here::here("data", "bq_stats.csv"), show_col_types = FALSE)

total_datasets <- nrow(stats)
total_tables <- sum(stats$n_tables)
total_gb <- sum(stats$total_bytes) / (1024^3)
n_projects <- n_distinct(stats$project)
```

```{r}
#| echo: false
layout_columns(
  value_box(
    title = "BigQuery Projects",
    value = n_projects,
    theme = value_box_theme(bg = "#fafafa", fg = "#1a1a1a"),
    style = "border-left: 2px solid #4A90D9;"
  ),
  value_box(
    title = "Datasets",
    value = total_datasets,
    theme = value_box_theme(bg = "#fafafa", fg = "#1a1a1a"),
    style = "border-left: 2px solid #4A90D9;"
  ),
  value_box(
    title = "Total Size",
    value = paste0(format(round(total_gb), big.mark = " "), " GB"),
    theme = value_box_theme(bg = "#fafafa", fg = "#1a1a1a"),
    style = "border-left: 2px solid #4A90D9;"
  ),
  col_widths = c(4, 4, 4)
)
```

The community currently brings together datasets from the following contributing projects:

- **[CWTS Leiden Datasets](cwts)** -- The CWTS data science team at Leiden University provides the CWTS Leiden Ranking Open Edition, time-specific OpenAlex versions, and other resources.
- **[InSySPo Campinas Datasets](insyspo/)** -- The InSySPo team at the University of Campinas provides time-specific versions of OpenAIRE and OpenAlex, among other resources.
- **[MultiObs](multiobs/)** -- These datasets from the University of Campinas support research on Science, Technology, and Innovation (STI) monitoring. The project develops a federated data infrastructure using participatory, multi-perspective approaches, including data on research entities, bibliometrics, economics, and intellectual property.
- **[Sesame Open Science Datasets (SOS)](sos/)** -- Sesame Open Science gives access to the latest public data snapshots of both Crossref and DataCite and the full OpenAIRE Graph Dataset. Other sources include ROR, DOAJ, and PKP. 
- **[SUB Göttingen Datasets](subugoe/)** -- The Scholarly Communication Analytics team at SUB Göttingen maintains monthly Crossref snapshots, OpenAlex releases, and other sources including Semantic Scholar and Unpaywall.

Want to share your data? Check our [contributing guide](../contributing.html).
