---
title: "Crosswalking ORION resources to measure open access to biomedical research literature"
author:
  - name: Najko Jahn
    orcid: 0000-0001-5105-1463
  - name: Bianca Kramer
    orcid: 0000-0002-5965-6560
  - name: Cameron Neylon
    orcid: 0000-0002-0068-716X
  - name: Nick Haupka
    orcid: 0009-0002-6478-6789
date: "2026-03-13"
abstract: "This tutorial demonstrates how to crosswalk various ORION resources. The focus is on funder and open access information for biomedical journal literature. Using the National Institutes of Health (NIH) and the European Commission (EC) as examples, the tutorial shows how to combine several data sources on BigQuery, including Crossref, OpenAIRE, OpenAlex, and open access agreement data from cOAlition S. The analysis covers more than 800,000 articles in PubMed, revealing notable differences in open access uptake between NIH- and EC-supported biomedical research."
categories: [tutorial, open access, funder info]
execute:
  cache: true
status: "wip"
---

```{r}
#| echo: false
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300,
  fig.retina = 2,
  dev = "ragg_png"
)

options(scipen = 999, digits = 1)
knitr::opts_chunk$set(format.args = list(big.mark = ",", digits = 2))
```

## Introduction

This notebook shows how several open research information resources provided by the [ORION community](https://orion-dbs.community/) on Google BigQuery can be used to measure open access by funder over time.
We use a comparison of NIH- and European Commission-supported biomedical research literature to illustrate the approach.
[Transformative agreements have played an important role in Europe](https://doi.org/10.1162/qss_a_00348) for several years, and [recent discussions around the NIH open access policy and price caps for publication fees](https://www.science.org/content/article/nih-s-proposed-caps-open-access-publishing-fees-roil-scientific-community) raise the question about the role of similar agreements in the US.

## Overview of data sources

Using this policy debate as a starting point, this notebook combines the following sources into a unified dataset on open access in biomedical research literature, with a particular focus on publisher-provided open access through transformative agreements.

*Identifying grant-supported biomedical literature*

- **[PubMed](https://pubmed.ncbi.nlm.nih.gov/)**: Our starting point, which is used to retrieve [NIH grant-supported publications](https://pubmed.ncbi.nlm.nih.gov/?term=NIH%5BGrants+and+Funding%5D) and as a standard source for biomedical literature. PubMed is not available through one of ORION's data offerings, so data are retrieved via the [Entrez API](https://www.ncbi.nlm.nih.gov/home/develop/api/). PubMed identifiers (PMIDs) serve as the linking key for biomedical research records.
::: {.column-margin}
The script used can be found [here](fetch_pmid_nih.R).
:::
- **[OpenAIRE](https://www.openaire.eu/)**: The [OpenAIRE graph](https://www.openaire.eu/), provided by [Sesame Open Science](https://orion-dbs.community/collections/sos/#openaire), is used to identify publications supported by the European Commission. OpenAIRE also contains PMIDs for PubMed-indexed articles, which allows filtering of to biomedical literature and linking to the other sources.

*Publication metadata*

- **[OpenAlex](https://openalex.org/)**: The primary source for bibliographic metadata in this tutorial. OpenAlex records also include PMIDs as identifiers. Here, we will use the [SUB Göttingen OpenAlex offering](https://orion-dbs.community/collections/subugoe/#openalex_walden).
- **[Crossref](https://www.crossref.org/)**: As provided by [SUB Göttingen](https://orion-dbs.community/collections/subugoe/#cr_instant). Useful for publisher disambiguation, because Crossref provides more consistent publisher names than OpenAlex.
- **Document classification**: A document classification dataset, developed as part of the [German Competence Network for Bibliometrics](https://open-bibliometrics.de/posts/20250717-DocumentTypeClassifier/) and provided by SUB Göttingen, is used to restrict the sample to original research articles.

*Open access information*

- **[OpenAlex](https://openalex.org/)**: Provides article- and journal-level open access status.
- **[cOAlition S Journal Checker Tool data](https://journalcheckertool.org/transformative-agreements/)**: Contains institution- and journal-level data crowd-sourced from library consortia. [SUB Göttingen consolidates this information and links articles to agreements](https://open-bibliometrics.de/posts/20250509-tarelease/). This tutorial shows how that matching can be carried out in Google BigQuery.

## Data gathering steps

### Setup

As a notebook software, we use [Quarto](https://quarto.org/). Acutally, the ORION website is also build with Quarto.

To query BigQuery from within our Quarto notebook, we use the R [`bigrquery`](https://bigrquery.r-dbi.org/) package via the DBI interface. The connection is established as follows:

```{r}
#| label: setup
#| message: false
library(tidyverse)
library(bigrquery)
library(DBI)
library(countrycode)

# Connect to GBQ with billing project from SUB Göttingen,
# You have to use your own :-)
bq_con <- dbConnect(
  bigrquery::bigquery(),
  project = "subugoe-collaborative",
  dataset = "openalex",
  billing = "subugoe-collaborative"
)
```

We also define a custom plot theme, inspired by the style used by [Our World in Data](https://ourworldindata.org/).

```{r}
my_theme <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.y = element_line(color = "grey85", linewidth = 0.3),
      axis.line.x = element_line(color = "grey40", linewidth = 0.4),
      axis.ticks.x = element_line(color = "grey40", linewidth = 0.3),
      axis.ticks.length.x = unit(0.15, "cm"),
      axis.ticks.y = element_blank(),
      axis.text = element_text(color = "grey30", size = rel(0.9)),
      axis.title.y = element_text(color = "grey30", size = rel(0.95), margin = margin(r = 10)),
      plot.title = element_text(face = "bold", size = rel(1.1), color = "grey10", margin = margin(b = 8)),
      plot.title.position = "plot",
      plot.caption = element_text(hjust = 1, size = rel(0.7), color = "grey50", margin = margin(t = 12)),
      plot.caption.position = "plot",
      legend.position = "top",
      plot.margin = margin(10, 10, 10, 10)
    )
}
```

### Create a publication metadata set

This section describes how we assembled the analytical dataset, combining grant funding information from PubMed and OpenAIRE with publication metadata from OpenAlex and Crossref.

It is generally good practice to begin by creating a focused subset of publications that contains only the columns needed for your analysis.
This reduces query costs, keeps the working dataset to a manageable size, and keeps a local subset of the raw metadata save.

Our first step is to build a publication dataset enriched with grant funding information from the NIH and the European Commission.
Because PubMed is not yet part of ORION, NIH funding data are retrieved directly from PubMed via the Entrez API.

The result is a list of PMIDs representing NIH-supported publications, which we then upload to BigQuery.

::: {.column-margin}
Here is the [R script](fetch_pmid_nih.R) and the resulting [PMID list](nih_pmids_2018_2024.txt) used.
:::

```{r}
pmids_nih <- read_lines("nih_pmids_2018_2024.txt")
dbWriteTable(bq_con, # Our BQ connections
             "subugoe-collaborative.resources.nih_pmids_2018_2024", # Table name
             tibble::tibble(pmids_nih = pmids_nih), # PMIDs in rectangular format with column name
             overwrite = TRUE
             )
```

In total, PubMed indexed `r length(unique(pmids_nih))` publications with NIH support between 2018 and 2024.

Next, we combine this list with OpenAIRE, [the most comprehensive source for EC-funded publications](https://arxiv.org/pdf/2109.10638),  to add European Commission-supported biomedical literature, and with OpenAlex and Crossref to obtain full publication metadata.

The resulting table is stored as `subugoe-collaborative.resources.oa_tutorial_md_raw`, and is publicly accessible. You will need to substitute your own dataset name.

```{sql}
#| connection: "bq_con"

CREATE OR REPLACE TABLE `subugoe-collaborative.resources.oa_tutorial_md_raw` AS (

WITH openaire_funders AS (
  SELECT DISTINCT
    pid.value AS pmid,
    fund.name AS funder
  FROM `sos-datasources.openaire.project_20260129` AS proj,
    UNNEST(fundings) AS fund
  INNER JOIN `sos-datasources.openaire.relation_product_project_20260129` AS pub_proj
    ON proj.id = pub_proj.target
  INNER JOIN `sos-datasources.openaire.publication_20260129` AS pub
    ON pub_proj.source = pub.id,
    UNNEST(pub.pids) AS pid
  WHERE fund.name IN ('European Commission')
    AND pid.scheme = 'pmid'
),

nih_funders AS (
  SELECT DISTINCT
    CAST(pmids_nih AS STRING) AS pmid
  FROM `subugoe-collaborative.resources.nih_pmids_2018_2024`
),

all_funders AS (
  SELECT pmid, 'European Commission' AS funder
  FROM openaire_funders
  UNION ALL
  SELECT pmid, 'NIH' AS funder
  FROM nih_funders
)

SELECT DISTINCT
    oalex.id,
    oalex.doi,
    oalex.ids.pmid,
    oalex.ids.pmcid,
    publication_year,
    publication_date,
    primary_location.source.issn_l AS issn_l,
    oalex.primary_location.source.host_organization_name AS oalex_publisher,
    cr.publisher AS cr_publisher,
    primary_location.source.is_in_doaj AS is_in_doaj,
    open_access.oa_status AS oa_status,
    country AS country_code,
    inst.ror,
    af.funder AS funders
  FROM `subugoe-collaborative.openalex_walden.works` AS oalex
  LEFT JOIN
    UNNEST(authorships) AS au WITH OFFSET AS pos
  LEFT JOIN
    UNNEST(au.countries) AS country
  LEFT JOIN
    UNNEST(au.institutions) AS inst
  LEFT JOIN `subugoe-collaborative.resources.document_classification_september25` AS doctype_classifier
    ON oalex.doi = doctype_classifier.doi
  INNER JOIN `subugoe-collaborative.cr_instant.snapshot` AS cr
    ON LOWER(oalex.doi) = LOWER(cr.doi)
  LEFT JOIN all_funders AS af
    ON REGEXP_EXTRACT(oalex.ids.pmid, r'(\d+)$') = af.pmid
  WHERE
      pos = 0
    AND primary_location.source.type = "journal"
    AND is_paratext = FALSE
    AND oalex.type IN ('article','review')
    AND is_research IS NOT FALSE
    AND is_xpac = FALSE
    AND (
      NOT REGEXP_CONTAINS(oalex.biblio.issue, '^[a-zA-Z]')
      OR oalex.biblio.issue IS NULL)
    AND (
      NOT REGEXP_CONTAINS(oalex.title, '[0-9]{3} pp.'))
    AND publication_year BETWEEN 2018 AND 2025
)
```


The query uses Common Table Expressions (CTEs), introduced with `WITH`, to build up the dataset in stages.
The first CTE, `openaire_funders`, extracts EC-funded publications from OpenAIRE by unnesting the fundings array, filtering for the European Commission, and joining to the OpenAIRE publication and relation tables to retrieve the associated PMIDs.
The second CTE, `nih_funders`, reads the PMID list we uploaded in the previous step.
Both are combined in `all_funders` using `UNION ALL`.

The main query retrieves metadata for all journal articles in the OpenAlex snapshot and joins the funder information.
Several filters are applied to keep the dataset focused on relevant publications: only journal articles and reviews are included; first author affiliations are used (identified via `WITH OFFSET` because the author position field was not available in this snapshot); and a set of pattern-based exclusions removes conference proceedings and other non-standard records.
Crossref is used as the source for publisher names, as it provides more consistent disambiguation than OpenAlex, which can split the same publisher across multiple imprints.
The dataset covers publications from 2018 to 2025.

### Estimate articles covered by transformative agreements

Next, we estimate how many open access articles were enabled by a transformative agreement, following the method described by [Jahn (2025)](https://link.springer.com/article/10.1007/s11192-025-05390-3).
Because of a lack of open access invoicing data, ususally not shared by libraries and publishers, various metadata sources ar ematched to obtain an estiamtion. Such an appraoch [has been proven to provide](https://doi.org/10.1162/QSS.a.24) robust results, but will likely underestimate the overall number of articles enabled by an angreements.

The query draws on several datasets from the Journal Checker Tool, provided by SUB Göttingen as part of the open bibliometric data offerings from the German Competence Network for Bibliometrics.

```{sql}
#| connection: "bq_con"
CREATE OR REPLACE TABLE `subugoe-collaborative.resources.oa_tutorial_jct_articles` AS (
WITH
  -- Enrich ROR variants from associated institutions
  obtain_associated_ror_ids AS (
    SELECT
      esac_id,
      jct_inst.ror_id AS ror_jct,
      inst.ror AS ror_associated
    FROM
      `subugoe-collaborative.openbib.jct_institutions` AS jct_inst
    LEFT JOIN
      `subugoe-collaborative.openalex.institutions` AS oalex_inst
    ON
      jct_inst.ror_id = oalex_inst.ror
    LEFT JOIN
      UNNEST(oalex_inst.associated_institutions) AS inst
  ),

  create_matching_table AS (
    SELECT esac_id, 'ror_jct' AS ror_type, ror_jct AS ror
    FROM obtain_associated_ror_ids
    UNION ALL
    SELECT esac_id, 'ror_associated' AS ror_type, ror_associated AS ror
    FROM obtain_associated_ror_ids
  ),

  enriched_ror_variants AS (
    SELECT DISTINCT
      create_matching_table.*,
      DATE(jct_inst.start_date) AS start_date,
      DATE(jct_inst.end_date) AS end_date
    FROM create_matching_table
    INNER JOIN `subugoe-collaborative.openbib.jct_esac` AS jct_inst
      ON create_matching_table.esac_id = jct_inst.id
  ),

  journal_agreements AS (
    SELECT
      j.issn_l,
      e.id AS esac_id
    FROM `subugoe-collaborative.openbib.jct_journals` AS j
    INNER JOIN `subugoe-collaborative.openbib.jct_esac` AS e
      ON j.esac_id = e.id
  )

SELECT DISTINCT
  md.id,
  md.doi,
  md.issn_l AS matching_issn_l,
  md.ror AS matching_ror,
  erv.ror_type,
  erv.esac_id,
  erv.start_date,
  erv.end_date,
  md.publication_date
FROM `subugoe-collaborative.resources.oa_tutorial_md_raw` AS md
INNER JOIN journal_agreements AS ja
  ON md.issn_l = ja.issn_l
INNER JOIN enriched_ror_variants AS erv
  ON md.ror = erv.ror
  AND ja.esac_id = erv.esac_id
WHERE
  md.ror IS NOT NULL
  AND md.issn_l IS NOT NULL
  AND (PARSE_DATE('%Y-%m-%d', md.publication_date) >= erv.start_date OR erv.start_date IS NULL)
  AND (PARSE_DATE('%Y-%m-%d', md.publication_date) <= erv.end_date OR erv.end_date IS NULL)
  -- Only TA OA articles
  AND oa_status IN ('gold', 'hybrid')
ORDER BY erv.esac_id, md.publication_date
)
```

With these two datasets in place, we are ready to proceed with the funder-specific open access analysis.

### Create the analytical dataset: funder view

The following query combines the publication metadata table with the transformative agreement articles table to count articles by funder, publication year, and open access type.
It also calculates the total number of articles per funder and year, which serves as the denominator for open access proportions.

The analysis focuses on articles made freely available on the journal website, covering gold open access in DOAJ-listed journals, hybrid open access, and diamond open access.
Repository-based open access, such as deposits in PubMed Central, is not distinguished further and is grouped into the "Other" category.
The same accounts for bronze, many biomedical journals make back issues openly available after an embargo period to comply with NIH open access policies prior 2025, which OpenAlex tags as bronze.

```{sql}
#| connection: "bq_con"
#| output.var: funding_oa
WITH
  oa AS (
    SELECT
      publication_year,
      funders,
      CASE
        WHEN (is_in_doaj = TRUE AND oa_status = 'gold') THEN 'gold_is_in_doaj'
        WHEN (is_in_doaj = TRUE AND oa_status = 'diamond')
          THEN 'diamond_is_in_doaj'
        WHEN oa_status = 'hybrid' THEN 'hybrid_oa'
        WHEN oa_status != 'closed' THEN 'other_oa'
        END AS oa_type,
      CASE
        WHEN jct.doi IS NOT NULL THEN TRUE
        ELSE FALSE
        END AS enabled_ta,
      COUNT(DISTINCT md.pmid) AS n_oa_articles
    FROM `subugoe-collaborative.resources.oa_tutorial_md_raw` AS md
    LEFT JOIN
      (
        SELECT DISTINCT doi
        FROM `subugoe-closed.apc_science.jct_articles`
      ) AS jct
      USING (doi)
    WHERE funders IS NOT NULL
    GROUP BY
      publication_year,
      funders,
      oa_type,
      enabled_ta
  ),
  total AS (
    SELECT
      publication_year,
      funders,
      COUNT(DISTINCT pmid) AS articles
    FROM `subugoe-collaborative.resources.oa_tutorial_md_raw`
    GROUP BY
      publication_year,
      funders
  )
SELECT
  total.publication_year,
  total.funders,
  total.articles,
  oa.n_oa_articles,
  oa.oa_type,
  oa.enabled_ta
FROM total
LEFT JOIN oa
  ON
    total.publication_year = oa.publication_year
    AND total.funders = oa.funders
WHERE total.funders IS NOT NULL AND total.publication_year < 2025
ORDER BY total.publication_year DESC
```

The query result is stored in an R object called `funding_oa` for further local analysis.

```{r}
funding_oa
```

```{r}
oa_by_year <- funding_oa |>
  filter(!is.na(oa_type)) |>
  summarise(n_oa = sum(n_oa_articles),
            .by = c(publication_year, funders, articles)
            ) |>
  mutate(oa_prop = n_oa / articles)
```

The following figure illustrates the relative growth of open access, comparing biomedical research articles acknowledging support from the European Commission and the NIH.
It is evident that both funders have similar open access shares of over 90%.

```{r}
#| dpi: 300
#| fig-retina: 2
#| label: oa_by_year_rel
oa_by_year |>
  ggplot(aes(publication_year, oa_prop, group = funders, color = funders)) +
  geom_line() +
  geom_point() +
  scale_y_continuous("Open Access (in%)", labels = scales::percent_format()) +
  scale_color_manual(NULL,
    values = c(
      "NIH" = "#323a45",
      "European Commission" = "#F8AE21"
    )) +
  labs(x = NULL,
       title = "Open Access to biomedical research literature",
       subtitle = "Journal articles and reviews in PubMed",
       caption = "Data sources: PubMed, OpenAIRE, OpenAlex.") +
  my_theme()
```

The next figure shows the absolute development.
```{r}
#| dpi: 300
#| fig-retina: 2
#| label: oa_by_year_abs
oa_by_year |>
  ggplot(aes(publication_year, n_oa, group = funders, color = funders)) +
  geom_line() +
  geom_point() +
  scale_y_continuous("Open Access", labels = scales::label_number(big.mark = ","), limits = c(0, NA), expand = c(0, 0)) +
  scale_color_manual(NULL,
    values = c(
      "NIH" = "#323a45",
      "European Commission" = "#F8AE21"
    )) +
  labs(x = NULL,
       title = "Open Access to biomedical research literature",
       subtitle = "Journal articles and reviews in PubMed",
       caption = "Data sources: PubMed, OpenAIRE, OpenAlex.") +
  my_theme()
```

Next, we break down open access by business model. Four categories are distinguished using a combination of OpenAlex data and estimates derived from the Coalition S Transformative Agreement data:

- **APC**: gold or hybrid open access, not via a transformative agreement
- **Transformative Agreement**: gold or hybrid open access enabled by a transformative agreement
- **Diamond OA**: diamond open access (no article processing charge)
- **Other**: all remaining open access routes (e.g. via PubMed Central)

Note that the focus is on publisher open access business models, and [OpenAlex open access status definition favors publisher-provided open access](https://help.openalex.org/hc/en-us/articles/24347035046295-Open-Access-OA), so articles in PMC repositories are accounted when there was no open version on a publisher website.

```{r}
oa_by_year_and_type <- funding_oa |>
  filter(!is.na(oa_type)) |>
  mutate(oa_cat = case_when(
    oa_type %in% c("gold_is_in_doaj", "hybrid_oa") & enabled_ta == FALSE ~ "APC",
    oa_type %in% c("gold_is_in_doaj", "hybrid_oa") & enabled_ta == TRUE ~ "Transformative Agreement",
    oa_type == "diamond_is_in_doaj" ~ "Diamond OA",
    .default = "Other")
) |> summarise(n_oa = sum(n_oa_articles),
            .by = c(publication_year, funders, articles, oa_cat)
            ) |>
  mutate(oa_prop = n_oa / articles,
             oa_cat = factor(oa_cat, levels = c("APC", "Transformative Agreement", "Diamond OA", "Other")))
```


```{r}
oa_by_year_and_type
```

The following two plots show the same breakdown as proportions and as absolute counts, respectively.

```{r}
#| fig-width: 8
#| dpi: 300
#| fig-retina: 2
#| label: oa_by_year_type_rel
oa_by_year_and_type |>
  ggplot(aes(publication_year, oa_prop, fill = fct_rev(oa_cat))) +
  geom_area(color = "black", linewidth = 0.2) +
  facet_wrap(~funders) +
  scale_y_continuous("Open Access (in%)", labels = scales::percent_format(), limits = c(0, 1),
                   expand = c(0, 0)) +
  scale_fill_manual("OA model",
    values = c(
      "APC" = "#0E6BA0",
      "Transformative Agreement" = "#30C0D2",
      "Diamond OA" = "#E2A3A4",
      "Other" = "#AEACA0"
    )) +
  labs(x = NULL,
       title = "Open Access to grant-supported biomedical research literature",
       subtitle = "Journal articles and reviews in PubMed",
       caption = "Data sources: PubMed, OpenAIRE, OpenAlex, cOAlition S Journal Checker Tool.") +
  my_theme() +
  guides(fill = guide_legend(reverse = TRUE))

```

```{r}
#| fig-width: 8
#| dpi: 300
#| fig-retina: 2
#| label: oa_by_year_type_abs

oa_by_year_and_type |>
  ggplot(aes(publication_year, n_oa, fill = fct_rev(oa_cat))) +
  geom_area(color = "black", linewidth = 0.2) +
  facet_wrap(~funders) +
  scale_y_continuous("Open Access", labels = scales::label_number(big.mark = ","),
                     limits = c(0, NA), expand = c(0, 0)) +
  scale_fill_manual("OA model",
    values = c(
      "APC" = "#0E6BA0",
      "Transformative Agreement" = "#30C0D2",
      "Diamond OA" = "#E2A3A4",
      "Other" = "#AEACA0"
    )) +
  labs(x = NULL,
       title = "Open Access to grant-supported biomedical research literature",
       subtitle = "Journal articles and reviews in PubMed",
       caption = "Data sources: PubMed, OpenAIRE, OpenAlex, cOAlition S Journal Checker Tool.") +
  my_theme() +
  guides(fill = guide_legend(reverse = TRUE))

```

```{r}
#| echo: false
options(scipen = 999, digits = 1)
knitr::opts_chunk$set(format.args = list(big.mark = ",", digits = 2))
```

The figures highlight differences in how open access is financed across the two funder communities.
In 2024, `r oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024) |> pull(oa_prop) |> sum() *100`% of NIH-supported biomedical research articles were open access (`r oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024) |> pull(n_oa) |> sum()` out of `r oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024) |> pull(articles) |> unique()` original research articles and reviews indexed in both PubMed and OpenAlex).
Of those open access articles, `r oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024, oa_cat == "APC") |> pull(n_oa) / oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024) |> pull(n_oa) |> sum()* 100`% were enabled through individual article processing charges (APCs), while only `r oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024, oa_cat == "Transformative Agreement") |> pull(n_oa) / oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024) |> pull(n_oa) |> sum() * 100`% were covered by transformative agreements.
The remaining `r oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024, oa_cat == "Other") |> pull(n_oa) / oa_by_year_and_type |> filter(funders == "NIH", publication_year == 2024) |> pull(n_oa) |> sum() * 100`% were made available through other routes, primarily via PubMed Central.

The picture is different for European Commission-funded research.
In 2024, `r oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024) |> pull(oa_prop) |> sum() *100`% of supported biomedical research articles were open access (`r oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024) |> pull(n_oa) |> sum()` out of `r oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024) |> pull(articles) |> unique()`).
Of those, `r oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024, oa_cat == "Transformative Agreement") |> pull(n_oa) / oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024) |> pull(n_oa) |> sum() * 100`% were covered by transformative agreements, a substantially higher share than for NIH-funded research.
This reflectsthe wider adoption of such institutional deals in Europe.
A further `r oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024, oa_cat == "APC") |> pull(n_oa) / oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024) |> pull(n_oa) |> sum()* 100`% were enabled through APCs, and `r oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024, oa_cat == "Other") |> pull(n_oa) / oa_by_year_and_type |> filter(funders == "European Commission", publication_year == 2024) |> pull(n_oa) |> sum() * 100`% through other routes.

The comparison suggests that agreements between library consortia and publishers play an important role in making European Commission-funded publications available in Europe, raising important questions about who pays for complying with the funder's open access policy.
The NIH in the US did achieve a similar overall open access share, but through a different route: a much larger proportion of articles is exclusively in the "Other" category, likely representing articles deposited in PubMed Central or made freely available on the publisher's website after an embargo period to comply with previous NIH mandates.
